Respiratory Disease Outbreaks in Toronto Pre, During, and Post Covid-19 Pandemic 

### 2025-05-13 paper/paper.qmd
very rough draft of discussion

@@

It becomes evident that since the end of the pandemic, the proportion of Coronavirus outbreaks compared to non-coronavirus outbreaks has been steadily decreasing.
## Non-Covid Respiratory Outbreaks
```{python}
#| label: no-corona-graph
#| fig-cap: "A line plot on a polar axis"

@@

## Rise in Non-Covid Respiratory Outbreaks {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Decrease in Covid Percentage of Respiratory Outbreaks

Please don't use these as sub-heading labels - change them to be what your point actually is.

@@

Weaknesses and next steps should also be included.

>>>>

Healthcare facilities include ....

@@

It becomes evident that since the end of the pandemic, the proportion of Coronavirus outbreaks compared to non-coronavirus outbreaks has been steadily decreasing.

## Non-Covid Respiratory Outbreaks
```{python}
#| label: no-corona-graph
#| fig-cap: "A line plot on a polar axis"

@@

It becomes apparent that given coronaviruses continue to comprise the greatest proportion of respiratory disease outbreaks in healthcare facilities in Toronto, and coronavirus outbreak trends match those of total respiratory outbreaks from, and total outbreaks from  2019-2024 that coronavirus outbreaks remain the biggest driver of disease outbreaks in Toronto. However, since the pandemic we have seen a decrease in both coronavirus cases, and percentage composition of all agents, indicating that its impact, while still pertinent, is decreasing.

Simultaneously, other respiratory infections including Influenza, Parainfluenza, RSV, Metapneumovirus, and Rhinovirus have been experiencing a significant increase in outbreak cases. Though respiratory infections such as Influenza are well known to the public, those such as RSV are known to a much lesser extent with only 40% of adults having heard of RSV, and 30% who would definitely get the RSV vaccine. Compared to Influenza, this is a much lesser. (https://www.canada.ca/en/public-health/services/immunization-vaccines/vaccination-coverage/seasonal-influenza-survey-results-2023-2024.html). It is important to note that while these respiratory infections are less deadly than covid (https://www.nature.com/articles/s41598-024-55378-x ), many of these healthcare facilities are long term healthcare facilities, whose populations are more susceptible to these viruses.(https://www.cdc.gov/respiratory-viruses/risk-factors/older-adults.html#:~:text=As%20people%20get%20older%2C%20their,increasing%20sharply%20with%20advancing%20age. )

Altogether, this suggests a slightly greater emphasis should be placed on non-covid respiratory diseases, specifically RSV given it is the only one of the above which has a vaccine available which is not widely known, and outbreak numbers have risen in recent years post-pandemic. By campaigning for increased RSV vaccinations, when outbreaks do occur in healthcare facilities, they will be less deadly.

@@

The main limitation of this study is the scale. All data is at the scale of healthcare facilities, thus must remain in the context of such. Extension of these results to the general Toronto population may be misrepresentative, thus to confirm the same trends hold in the general public further research should be conducted with data coming from the general population.
